---
title: "miRNA_Target_Finder"
author: "Nick Werry"
date: "11/12/2020"
output: html_document
params:
  input_file:
    label: "load input file OR enter sequences below"
    input: file
    value: "Reem_File.xlsx"
  miRNA_A:
    label: "first miRNA of interest (hsa-miR-##-#p)"
    input: text
    value: ""
  miRNA_B:
    label: "second miRNA of interest (hsa-miR-##-#p)"
    input: text
    value: ""
  valResults: 
    label: "Validated interactions"
    value: TRUE
  preResults:
    label: "Predicted interactions" 
    value: TRUE
  numResults:
    label: "How many targets to return?"
    input: text
    value: 10
  saveFile:
    label: "Save output file of all data?"
    value: TRUE
  numDatabases:
    label: "ADVANCED: How many databases must predict?"
    input: slider
    min: 1
    max: 8
    step: 1
    value: 4
  Beep:
    label: "Beep when finished"
    value: FALSE
---

<!-- USE this line in the console to knit with custom parameters -->
<!-- rmarkdown::render("miRNA_Target_Finder", params = "ask") -->

```{r Setup, echo = FALSE}
library(knitr)

#Tidy output
opts_chunk$set(tidy = TRUE, cache=FALSE)

#Limit decimals
options(digits = 3)

#Format kables
options(knitr.table.format = 'markdown')
```

```{r Libraries, echo = FALSE, message = FALSE, warning = FALSE}
if ("tidyverse" %in% rownames(installed.packages()) == FALSE) {install.packages("tidyverse")}
library(tidyverse)
library(writexl)
library(readxl)
library(ggplot2)
if ("BiocManager" %in% rownames(installed.packages()) == FALSE) {install.packages("BiocManager")}
library(BiocManager)
if ("multiMiR" %in% rownames(installed.packages()) == FALSE) {BiocManager::install("multiMiR")}
library(multiMiR)
if ("beepr" %in% rownames(installed.packages()) == FALSE) {install.packages("beepr")}
library(beepr)
```


```{r Functions, echo = FALSE}
#Return formatted list of Query or Confounding miRNAs from input file, removing NAs
miRNA_find <- function(input_df, miRNA_set, row = 1) {
  #Select range of file depending on if miRNA is query or confounding (miRNA_set)
  if (miRNA_set == "query") {
    miR_list <- c(input_df[row, 1:2])
  } else if (miRNA_set == "confounding") {
    miR_list <- c(input_df[1, 3:12])
  }
  #Remove NAs, format to add hsa- if not already present
  miR_list_no_NA <- miR_list[!is.na(miR_list)]
  for (miR in miR_list_no_NA) {
    miR_list_no_NA[[miR]] <- ifelse(startsWith(miR, "hsa-"), miR, sprintf("hsa-%s", miR))
  }
  hsa_miR_list <- miR_list_no_NA[-(1:length(miR_list[!is.na(miR_list)]))]
  return(hsa_miR_list)
}

#Return tidied dataframe of miRNA targets
target_find <- function(mirna_list, database_type) {
  #Collect data using multiMiR
  all_data <- multiMiR::get_multimir(org = "hsa", mirna = mirna_list, table = database_type)
  #Process data if searching validated databases
  if (database_type == "validated") {
    target_df <- data.frame(all_data@data) %>%
    dplyr::select(3,4,9,10) 
  #Process data if searching predicted databases
  } else if (database_type == "predicted"){
    target_df <- data.frame(all_data@data) %>%
    dplyr::select(1,3,4,7,8)
  }
  #Tidy to remove irrelevant or redundant results
  target_clean <- target_df %>%
    filter(!(target_symbol == "")) %>%
    filter(type == database_type) %>%
    dplyr::select(-type) %>%
    distinct()
  return(target_clean)
}

#Add columns counting number of databases predicting interactions, filter for indv. minimum
database_count <- function(target_df) {
  count_df <- target_df %>%
  group_by(mature_mirna_id, target_symbol) %>%
  mutate(indv_database_count = length(unique(database))) %>%
  group_by(target_symbol) %>%
  mutate(cotarget_database_count = length(unique(database))) %>%
  distinct() 
}

#Return targets shared by all Query miRNAs
cotarget_find <- function(target_df){
  input_mirnas <- length(unique(target_df$mature_mirna_id))
  cotargets <- target_df %>%
    group_by(target_symbol) %>%
    mutate(miRNA_count = length(unique(mature_mirna_id))) %>%
    filter(miRNA_count == length(input_mirnas))
}

#Order results of cotargeting
cotarget_rank <- function(cotarget_df){
  #Process data if searching validated databases
  if ("pubmed_id" %in% colnames(cotarget_df)){
    ranked_result <- as.data.frame(table(cotarget_df$target_symbol)) %>%
      arrange(desc(Freq)) %>%
      rename("target_symbol"="Var1","reports"="Freq")
  #Process data if searching predicted databases
  } else if ("score" %in% colnames(cotarget_df)){
    ranked_result <- cotarget_df %>%
      group_by(mature_mirna_id) %>%
      group_by(database) %>%
      mutate(rank = order(order(database))) %>% #Follow default score ranking
      group_by(target_symbol) %>%
      mutate(mean_rank = mean(rank)) %>%
      arrange(mean_rank) %>%
      arrange(desc(cotarget_database_count)) %>%
      dplyr::select(target_symbol, cotarget_database_count, mean_rank) %>%
      distinct()
  }
}
```

```{r Parameters, echo = FALSE}
#Load inputs from file or text input
  #Load from text
if (params$input_file == "") {
  miRNA_query <- vector(mode = "list")
  miRNA_confounding <- vector(mode = "list") #Empty list to allow for ifs later on
  for (miR in params){
    if (str_detect(miR, "miR")){
      mirList[miR] <- (ifelse(startsWith(miR, "hsa-"), miR, sprintf("hsa-%s", miR)))
    }
  }
} else {
  #Load from file
  input_df <- read_excel(params$input_file)
  #Create lists of miRNAs of interest
  miRNA_query <- miRNA_find(input_df, "query")
  miRNA_confounding <- miRNA_find(input_df, "confounding")
}

#Create string of abbreviated miRNA inputs
query_short <- vector(mode = "character")
for (miR in miRNA_query){
  query_short[miR] <-  str_extract(miR, "(miR\\-)\\w+")
}
confounding_short <- vector(mode = "character")
for (miR in miRNA_confounding){
  confounding_short[miR] <-  str_extract(miR, "(miR\\-)\\w+")
}

#Visually appealing string of miRNA_short
q_amp <- paste(query_short, collapse = " & ")
c_amp <- paste(confounding_short, collapse = " & ")

#Variables for filenames
fileName <- sprintf("Output %s.xlsx", q_amp)

#Set blank dfs to allow for no confounding
confounding_predicted <- data.frame(target_symbol = NULL) 
confounding_validated <- data.frame(target_symbol = NULL) 
```
### `r q_amp` targets `r if (length(miRNA_confounding) != 0){sprintf("without overlap from %s", c_amp)}`\

#### `r if (length(miRNA_confounding) != 0) {sprintf("%s screened for confounding effects", c_amp)}`
```{r Confounding Targets, echo = F, eval = (length(miRNA_confounding) != 0), message = F}
#Create vector of predicted results
confounding_predicted <- target_find(miRNA_confounding, "predicted") %>%
  dplyr::select(target_symbol) %>%
  distinct()

#Get vector of validated targets of confounding miRNAs
confounding_validated <- target_find(miRNA_confounding, "validated") %>%
  dplyr::select(target_symbol) %>%
  distinct()
```
`r sprintf("%s predicted and %s validated confounding interactions", length(confounding_predicted$target_symbol), length(confounding_validated$target_symbol))`

#### `r if(params$valResult){sprintf("Top %s Validated Results", params$numResults)}`
```{r Validated, echo = F, eval = params$valResults}
#Get validated results for query miRNAs
query_validated <- target_find(miRNA_query, "validated")
qv_cotarget <- cotarget_find(query_validated)
qv_rank <- cotarget_rank(qv_cotarget)

#Filter validated results to remove targets shared by confounding miRNAs
qv_no_cv <- qv_rank %>%
  filter(!(target_symbol %in% confounding_validated$target_symbol))

#Reference against predicted results
qv_no_cp <- qv_rank %>%
  filter(!(target_symbol %in% confounding_predicted$target_symbol))

#Reference against both predicted and validated results
qv_no_cv_cp <- qv_no_cv %>%
  filter((target_symbol %in% qv_no_cp$target_symbol))

#Output number of removed targets
v_diff <- length(qv_rank$target_symbol) - length(qv_no_cv_cp$target_symbol)
cat(sprintf("%s targets shared by confounding miRNAs removed. Top %s of %s shown.", v_diff, params$numResults, length(qv_no_cv_cp$target_symbol)))

#Output top filtered result
qv_no_cv_cp %>%
  head(n = params$numResults) %>%
  kable(format = "simple")
```

#### `r if(params$preResult){sprintf("Top %s Predicted Results", params$numResults)}`
```{r Predicted, echo = F, eval = params$preResults}
#Get predicted results for query miRNAs
query_predicted <- target_find(miRNA_query, "predicted")
qp_databases <- database_count(query_predicted)
qp_databases_filt <-  qp_databases %>% filter(indv_database_count >= params$numDatabases) #Filter from input slider to avoid biasing
qp_cotarget <- cotarget_find(qp_databases_filt)
qp_rank <- cotarget_rank(qp_cotarget)

#Reference against predicted results
qp_no_cp <- qp_rank %>%
  filter(!(target_symbol %in% confounding_predicted$target_symbol))

#Reference against previous vector of validated results
qp_no_cv <- qp_rank %>%
  filter(!(target_symbol %in% confounding_validated$target_symbol))

#Reference against both predicted and validated results
qp_no_cv_cp <- qp_no_cp %>%
  filter((target_symbol %in% qp_no_cv$target_symbol))

#Plot database restriction to see how many interactions are being ignored
ggplot(qp_databases, mapping = aes(indv_database_count)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = (1:8)) +
  geom_vline(xintercept = (params$numDatabases - 0.5), colour = "red") +
  ggtitle(sprintf("Predicted results filtered for presence in %s or more databases", params$numDatabases))

#Output number of removed targets
p_diff <- length(qp_rank$target_symbol) - length(qp_no_cv_cp$target_symbol)
cat(sprintf("%s targets shared by confounding miRNAs removed. Top %s of %s shown.", p_diff, params$numResults, length(qp_no_cv_cp$target_symbol)))

#Output top result
qp_no_cv_cp %>%
  head(n = params$numResults) %>%
  kable(format = "simple")
```


```{r Save Results, echo = F, eval = params$saveFile}
#Run analyses if not performed above
if (params$valResults == FALSE){
query_validated <- target_find(miRNA_query, "validated")
qv_cotarget <- cotarget_find(query_validated)
qv_rank <- cotarget_rank(qv_cotarget)
confounding_validated <- target_find(miRNA_confounding, "validated") %>%
  dplyr::select(target_symbol) %>%
  distinct()
qv_no_cv <- qv_rank %>%
  filter(!(target_symbol %in% confounding_validated$target_symbol))
}
if (params$preResults == FALSE){
#Get predicted results for query miRNAs
query_predicted <- target_find(miRNA_query, "predicted")
qp_databases <- database_count(query_predicted) %>%
  filter(indv_database_count >= params$numDatabases)
qp_cotarget <- cotarget_find(qp_databases)
qp_rank <- cotarget_rank(qp_cotarget)
qp_no_cv <- qp_rank %>%
  filter(!(target_symbol %in% confounding_validated$target_symbol))
}

#List of data for excel
data_output <- if (length(miRNA_confounding) == 0) {
    list(Pred. = qp_rank, Val. = qv_rank)
} else {
    list(Pred. = qp_rank, Val. = qv_rank, Pred_no_CV = qp_no_cv, Val_no_CV = qv_no_cv, Pre_no_CP = qp_no_cp, Pre_no_CP_CV = qp_no_cv_cp, CV = confounding_validated, CP = confounding_predicted)
}

writexl::write_xlsx(x = data_output, path = fileName)
```

```{r Beep, echo = FALSE, eval = params$Beep}
beepr::beep(1)
Sys.sleep(.5)
beepr::beep(1)
Sys.sleep(.5)
beepr::beep(1)
```